---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r readin}
library(dplyr)
library(ggplot2)
data<- read.csv('activity.csv')
no_na_data<- tbl_df(data)
data<- tbl_df(data)
no_na_data[is.na(no_na_data)]<- 0
```

## What is mean total number of steps taken per day?
```{r average_steps_per_day}
total_step<- no_na_data%>%group_by(date)%>%summarise(sum(steps))
qplot(`sum(steps)`,data = total_step, geom = 'histogram',binwidth=500,xlab = 'total number of steps per day')
mean_total_step<- mean(total_step$'sum(steps)')
median_total_step<- median(total_step$'sum(steps)')
```
The mean of total number of steps taken per day is *`r as.character(mean_total_step)`*.  
The median of total number of steps taken per day is *`r as.character(median_total_step)`*

## What is the average daily activity pattern?
```{r average_daliy_activities}
data<- filter(data,!is.na(steps))
average_step<- data%>%group_by(interval)%>%summarise(mean(steps))
qplot(interval,`mean(steps)`,data = average_step,geom = 'line',xlab = '5 min intervals',ylab = 'average steps')
max_steps<- max(average_step[,'mean(steps)'])
max_ID<- average_step%>%filter(`mean(steps)`==max_steps)%>%select(interval)
```
Interval *`r max_ID`* on average across all the days in the dataset, contains the maximum number of steps

## Imputing missing values
```{r imputing_mission_value}
data<- read.csv('activity.csv')
new_data<- tbl_df(data)
na_data<- filter(new_data,is.na(steps))
na_rows<- nrow(na_data)
for (i in 1:nrow(new_data)){
  if (is.na(new_data[i,"steps"])){
    interval<- as.integer(new_data[i,'interval'])
    position<- average_step[,'interval']==interval
    new_data[i,'steps']<- average_step[position,2]
  }
}
total_step<- new_data%>%group_by(date)%>%summarise(sum(steps))
qplot(`sum(steps)`,data = total_step, geom = 'histogram',binwidth=1000,xlab = 'total number of steps per day')
mean_total_step<- mean(total_step$'sum(steps)')
median_total_step<- median(total_step$'sum(steps)')
```
The mean of total number of steps taken per day is *`r as.character(mean_total_step)`*.  
The median of total number of steps taken per day is *`r as.character(median_total_step)`*

## Are there differences in activity patterns between weekdays and weekends?
```{r weekday_vs_weekend_pattern}
new_data<- mutate(new_data,day= ifelse(weekdays(as.Date(date,'%Y-%m-%d'))%in%c('Saturday','Sunday'),'weekend','weekday'))
new_data<- group_by(new_data,interval,day)
average_step<- new_data%>%summarise(mean(steps))
qplot(interval,`mean(steps)`,data = average_step,geom='line',facets = day~.,xlab = '5 min intervals',ylab = 'average steps',main = 'Activity patterns between weekdays and weekends')
```
